Towards a Mechanized Proof of Noninterference

### 2013-12-18 paper/paper.tex
Ugh I'm tired, I keep typing -> instead of \to.
Outlined cases for proof of subject reduction.  Wrote an initial draft of the conclusion.

\begin{figure}[t]
  \begin{align*}
    t ::= int^l \: | \: (t -> t)^l
  \end{align*}
  \caption{Types for \langName}
  \label{fig:types}
\end{figure}

@@

\begin{lemma}[Subject Reduction]
  \label{lem:subj_red}
  Let $e \to e'$.  Assume $\typeRule{pc}{e}{t}$ and $pc \in H$ for all $i \in
  \{1, 2\}$.  Then $\typeRule{pc}{e'}{t}$.
\end{lemma}
\begin{proof}
  \todo{}
\end{proof}

@@

Finally, the noninteference theorem shows that if we have an expression that is
typed with a low label, then replacing any bracketed value in the expression
with either projection will result in the same value for the overall expression.
In other words, if we vary any high security item in a low security expression,
the value must not be affected.

\begin{thm}[Noninterference]
  Choose $l, h \in \mathcal{L}$ such that $h \not\leq l$.  Let $h \guards t$.
  Assume $\typeRule{(x \mapsto t)}{e}{\textsf{int}^l}$, where $e$is a \langName
  expression.  If, for all $i \in \{1, 2\}$, $\typeRule{ }{v_i}{t}$ and $e[x
  \Leftarrow v_i] \to^* v_i'$ hold, then $v_1' = v_2'$.
\end{thm}
\begin{proof}
  Let $H$ be the upward closure of $\{h\}$ and $v = < v_1 | v_2 >$, since other
  $v$ would be immediately uninteresting.  By \textsc{V-Bracket}, we know that
  $\typeRule{ }{v}{t}$ holds given the assumptions.
  By Lemma~\ref{lem:substitution}, we have $\typeRule{ }{e[x \Leftarrow
  v]}{\textsf{int}^l}$.  By Lemma~\ref{lem:final_support}, we know that
  $\lift{e[x \Leftarrow v]}_i = e[x \Leftarrow \lift{v}_i]$, which under the
  current assumptions reduces to $v_i$.  By Lemma~\ref{lem:completeness} we know
  that there is an answer $a$ such that $e[x \Leftarrow v] ->^* a$.  By
  Lemma~\ref{lem:soundness}, $\lift{a}_i = v_i'$ for all $i$ which means that
  $a$ is a value.  Then, we know by Lemma~\ref{lem:final_support} that $v_1' =
  v_2'$.
\end{proof}

@@

\todo{Summarize again what we covered, where we ended up, and where we'd like to
go next.}

>>>>

\begin{figure}[t]
  \begin{align*}
    t ::= int^l \: | \: (t \to t)^l
  \end{align*}
  \caption{Types for \langName}
  \label{fig:types}
\end{figure}

@@

\begin{lemma}[Subject Reduction]
  \label{lem:subj_red}
  Let $e \to e'$.  Assume $\typeRule{pc}{e}{t}$ and $pc \in H$ for all $i \in
  \{1, 2\}$.  Then $\typeRule{pc}{e'}{t}$.
\end{lemma}
\begin{proof}
  By induction on the derivation of $e \to e'$.
  \begin{case}[$\beta$]
    \todo{}
  \end{case}

  \begin{case}[if]
    \todo{}
  \end{case}

  \begin{case}[let]
    \todo{}
  \end{case}

  \begin{case}[lift-app]
    \todo{}
  \end{case}

  \begin{case}[lift-if]
    \todo{}
  \end{case}

  \begin{case}[bracket]
    \todo{}
  \end{case}
\end{proof}

@@

The noninteference theorem shows that if we have an expression that is typed
with a low label, then replacing any bracketed value in the expression with
either projection will result in the same value for the overall expression.  In
other words, if we vary any high security item in a low security expression, the
value must not be affected.

\begin{thm}[Noninterference]
  Choose $l, h \in \mathcal{L}$ such that $h \not\leq l$.  Let $h \guards t$.
  Assume $\typeRule{(x \mapsto t)}{e}{\textsf{int}^l}$, where $e$is a \langName
  expression.  If, for all $i \in \{1, 2\}$, $\typeRule{ }{v_i}{t}$ and $e[x
  \Leftarrow v_i] \to^* v_i'$ hold, then $v_1' = v_2'$.
\end{thm}
\begin{proof}
  Let $H$ be the upward closure of $\{h\}$ and $v = < v_1 | v_2 >$, since other
  $v$ would be immediately uninteresting.  By \textsc{V-Bracket}, we know that
  $\typeRule{ }{v}{t}$ holds given the assumptions.
  By Lemma~\ref{lem:substitution}, we have $\typeRule{ }{e[x \Leftarrow
  v]}{\textsf{int}^l}$.  By Lemma~\ref{lem:final_support}, we know that
  $\lift{e[x \Leftarrow v]}_i = e[x \Leftarrow \lift{v}_i]$, which under the
  current assumptions reduces to $v_i$.  By Lemma~\ref{lem:completeness} we know
  that there is an answer $a$ such that $e[x \Leftarrow v] \to^* a$.  By
  Lemma~\ref{lem:soundness}, $\lift{a}_i = v_i'$ for all $i$ which means that
  $a$ is a value.  Then, we know by Lemma~\ref{lem:final_support} that $v_1' =
  v_2'$.
\end{proof}

@@

In this paper, we present a simple information flow typed language \langName,
prove noninterference for the language, and discuss our experiences beginning to
mechanize this proof in coq.  Although we did not complete our proof, we were
able to catch inaccuracies and subtleties that the proof we based our approach
on did not handle or explicitly cover.  This is the first step towards a
verified compiler for a language with information flow types such as Jif or
FlowCaml, which allows for a more trustworthy compiler and language definition,
by allowing users to verify the proofs in a proof assistant like Coq.  As far as
we know, a verified proof of noninterference for such a language has not been
implemented, let alone a verified compiler.  We are confident that, given more
time and a large amount of work, such a verified compiler could be realized for
a mature information flow language.
